---
title: "HMIS exploration Uttar Pradesh"
author: "Dan Weinberger"
date: "October 29, 2019"
output: html_document
---

PCV was launched by the Union Health Minister, Shri JP Nadda on May 13, 2017 at Mandi,
Himachal Pradesh [1]. With this phased introduction, nearly 2.1 million children in
Himachal Pradesh (all 12 districts), parts of Bihar (17 out of 38 districts)
and Uttar Pradesh (6 out of 75 districts) will be vaccinated with PCV in the first year [2].
This will be followed by introduction in Madhya Pradesh and Rajasthan next year, and eventually coverage will be expanded across the entire country in a phased manner, in the coming years. "
In uttar Pradesh it is: Lakhimpur Kheri, Sitapur, Siddharth Nagar, Bahraich, Balrampur, Shrawasti;
In Bihar:  The 17 high-priority districts are  Araria, Begusarai, Darbhanga, Kishanganj, Khagaria, Katihar,
Muzaffarpur, Munger, Vaishali, Madhepura, Madhubani, Purnea, Samastipur, Saran, Sitamarhi, Sheohar and Supaul

#######
Lorine
ACUTE RESPIRATORY INFECTION REQUIRING ADMISSION??
85% of DEATH OCCUR AT HOME--SHOULD BE IN HMIS; MANY
IN PRIVATE SECTOR HOSPITALS 
CONTROLS: DIARRHEA deaths? ROTA VACCINE IS ROLLING OUT..ACUTE ENCAPHALITIS IN SOUTH UP AND BIHAR. THERE ARE SOME VACCINE PROGRAMS BUT STARTED 5 YEARS AGO...NOT REALLY IN WESTERN UP
ASPHYXIA DEATHS WONT BE SAME HOSPITALS --WOULD BE IN WOMENS HOSPITAL...MANY ARE SEPARATE WOMENS/MENS HOSPTALS
UP HMIS--DR VISANTH  
#######
 
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reshape2)

```

##Load data
```{r}
bh1<-read.csv('uttarPradesh series to Mar 2019.csv')

```

#Data structure
```{r}
str(bh1)

```

Relabel data 
```{r}
bh1$month<-as.numeric(substr(bh1$DATE,6,7))
bh1$year<-as.numeric(substr(bh1$DATE,2,5))
bh1$uri<-bh1$X_10_13
bh1$pneu_death<-bh1$X_16_3_1
bh1$diar_death<-bh1$X_16_3_2
bh1$measles_death<-bh1$X_16_3_4
bh1$asphyxia_death<-bh1$X_16_2_2
bh1$sepsis_death<-bh1$X_16_2_1
bh1$neonatal_death<-bh1$X_16_1
bh1$monthdate<-as.Date(paste(bh1$year,bh1$month,'01', sep='-'))
```

Flag the districts that have introduced PCV
```{r}
up.intro.districts<-c('Lakhimpur Kheri', 'Sitapur', 'Siddharth Nagar', 'Bahraich', 'Balrampur', 'Shrawasti')
bh1$pcv.status<-0
bh1$pcv.status[bh1$DISTRICT %in% up.intro.districts] <-1
```

Time series plots
```{r}

strat1<-factor(bh1$monthdate)
ds.sub<-bh1[,c('uri', 'diar_death', 'pneu_death','sepsis_death','asphyxia_death', 'measles_death', 'neonatal_death')]

#Aggregate to state-level
bh2<-aggregate(x=ds.sub, by=list( strat1) , FUN='sum', na.rm=TRUE) 
names(bh2)<-c('monthdate',names(ds.sub))
bh2$monthdate<-as.Date(bh2$monthdate)
bh2$neonatal_death[nrow(bh2)]<-NA
bh2$month<-as.factor(month(bh2$monthdate))
bh2$year<-as.factor(year(bh2$monthdate))
par(mfrow=c(1,1), mar=c(3,3,1,1))
plot(bh2$monthdate,bh2$pneu_death,main='Pneumonia deaths', type='l', bty='l')
```
Reshape data and make a heatmap by district
```{r}
n.cols<-100
nice.cols <-  colorRampPalette(brewer.pal(11, "YlOrRd"))(n.cols)


ds.sub<-bh1[,c('DISTRICT','monthdate','uri')]
ds.m<-melt(ds.sub, id=c('DISTRICT','monthdate'))
ds.c<-dcast(ds.m, monthdate~DISTRICT)
par(mfrow=c(1,1), mar=c(1,1,1,1))
hm1<-heatmap(t(as.matrix(ds.c[,-1])), scale='row', Rowv=NA, Colv=NA,cexRow =0.5, col=nice.cols)

```

